Plug-and-play dual-tree algorithm runtime analysis

### 2015-01-21 bounds.tex
Some changes.
Some minor last-minute changes.

\author{\name Ryan R. Curtin \email ryan@ratml.org\\
  \addr School of Computational Science and Engineering\\
  Georgia Institute of Technology\\
  Atlanta, GA 30332-0250, USA
  \AND
  \name Dongryeol Lee \email drselee@gmail.com\\
  \addr GE Research\\
  Schenedectadadaty, NY\\
  \AND
  \name William B. March \email march@ices.utexas.edu\\
  \addr Austin, TX, USA\\
  \AND
  \name Parikshit Ram \email p.ram@gatech.edu\\
  \addr Skytree
}

@@

\begin{abstract}
%A large collection of machine learning algorithms contain subproblems that can
%be efficiently solved with trees.  In the situation where a dual-tree algorithm
%with cover trees is used to solve a problem, rigorous worst-case runtime bounds
%can often be shown.  A worst-case runtime bound is proven for all cover tree
%dual-tree algorithms; this bound can be employed to give useful runtime bounds
%for individual problems.  We show a new, tighter bound of $O(c^9 \nu N)$ for
%all-nearest-neighbor search and also provide linear runtime bounds for
%approximate kernel density estimation and exact range count/search under some
%mild assumptions.  These are the first linear runtime bounds we are aware of for
%range count/search.
%Numerous machine learning algorithms contain pairwise statistical problems at
%their core. Often, tree structures are used to solve these problems efficiently.
%Dual-tree algorithms using cover trees have rigorous worst-case runtime
%guarantees for some of these problems. In this paper, we present a {\em
%problem-independent} runtime guarantee for {\em any} dual-tree algorithm using
%the cover tree, separating out the problem-dependent and the problem-independent
%elements. This allows us to just plug in bounds for the problem-dependent
%elements to get runtime guarantees for dual-tree algorithms for any pairwise
%statistical problem. We demonstrate this plug-and-play procedure for
%nearest-neighbor search and approximate kernel density estimation to get
%improved runtime guarantees.  Under some assumptions, we also present the first
%linear runtime guarantee for dual-tree based range search.
Numerous machine learning algorithms contain pairwise statistical problems at
their core -- i.e. tasks that require computations over all pairs of input points if implemented directly. 
Often, tree structures are used to solve these problems efficiently. Dual-tree algorithms 
can efficiently solve or approximate many of these problems.
Using cover trees, rigorous worst-case runtime
guarantees have been proven for some of these algorithms. In this paper, we present a {\em
problem-independent} runtime guarantee for {\em any} dual-tree algorithm using
the cover tree, separating out the problem-dependent and the problem-independent
elements.  This allows us to just plug in bounds for the problem-dependent
elements to get runtime guarantees for dual-tree algorithms for any pairwise
statistical problem without re-deriving the entire proof. We demonstrate this plug-and-play procedure for
nearest-neighbor search and approximate kernel density estimation to get
improved runtime guarantees.  Under mild assumptions, we also present the first
linear runtime guarantee for dual-tree based range search.
\end{abstract}

>>>>

\author{\name Ryan R. Curtin \email ryan@ratml.org\\
  \addr School of Computational Science and Engineering\\
  Georgia Institute of Technology\\
  Atlanta, GA 30332-0250, USA
  \AND
  \name Dongryeol Lee \email drselee@gmail.com\\
  \addr GE Global Research Center\\
  Schenectady, NY 12309
  \AND
  \name William B. March \email march@ices.utexas.edu\\
  \addr Institute for Computational Engineering and Sciences\\
  University of Texas, Austin\\
  Austin, TX 78712-1229
  \AND
  \name Parikshit Ram \email p.ram@gatech.edu\\
  \addr Skytree, Inc.\\
  Atlanta, GA 30332
}

@@

\begin{abstract}
%A large collection of machine learning algorithms contain subproblems that can
%be efficiently solved with trees.  In the situation where a dual-tree algorithm
%with cover trees is used to solve a problem, rigorous worst-case runtime bounds
%can often be shown.  A worst-case runtime bound is proven for all cover tree
%dual-tree algorithms; this bound can be employed to give useful runtime bounds
%for individual problems.  We show a new, tighter bound of $O(c^9 \nu N)$ for
%all-nearest-neighbor search and also provide linear runtime bounds for
%approximate kernel density estimation and exact range count/search under some
%mild assumptions.  These are the first linear runtime bounds we are aware of for
%range count/search.
%Numerous machine learning algorithms contain pairwise statistical problems at
%their core. Often, tree structures are used to solve these problems efficiently.
%Dual-tree algorithms using cover trees have rigorous worst-case runtime
%guarantees for some of these problems. In this paper, we present a {\em
%problem-independent} runtime guarantee for {\em any} dual-tree algorithm using
%the cover tree, separating out the problem-dependent and the problem-independent
%elements. This allows us to just plug in bounds for the problem-dependent
%elements to get runtime guarantees for dual-tree algorithms for any pairwise
%statistical problem. We demonstrate this plug-and-play procedure for
%nearest-neighbor search and approximate kernel density estimation to get
%improved runtime guarantees.  Under some assumptions, we also present the first
%linear runtime guarantee for dual-tree based range search.
Numerous machine learning algorithms contain pairwise statistical problems at
their core---that is, tasks that require computations over all pairs of input
points if implemented naively.  Often, tree structures are used to solve these
problems efficiently. Dual-tree algorithms can efficiently solve or approximate
many of these problems.  Using cover trees, rigorous worst-case runtime
guarantees have been proven for some of these algorithms. In this paper, we
present a {\em problem-independent} runtime guarantee for {\em any} dual-tree
algorithm using the cover tree, separating out the problem-dependent and the
problem-independent elements.  This allows us to just plug in bounds for the
problem-dependent elements to get runtime guarantees for dual-tree algorithms
for any pairwise statistical problem without re-deriving the entire proof. We
demonstrate this plug-and-play procedure for nearest-neighbor search and
approximate kernel density estimation to get improved runtime guarantees.  Under
mild assumptions, we also present the first linear runtime guarantee for
dual-tree based range search.
\end{abstract}
